Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-05 paper.tex
grammer reprhrese

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus of this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum. LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance. In this work, a simulation environment to study different SF assignment schemes is implemented. Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization. Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on growth of number of IoT applications. LPWAN technologies addresses some of the well-known wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications do not require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox are known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. LoRaWAN networks can be deployed as a private network. However, Sigfox and NB-IoT are only available with operator contracted deployments \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Also, Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

@@

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.

\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides inter-operability between different LoRa networks. LoRaWAN is based on pure ALOHA medium access. LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is low power embedded device that communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios. Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateway. Class C end nodes extend Class A by keeping receive window open unless they are transmitting. This provides low latency donwlink communication but requires more power consumption. In this paper, we focused on Class A end devices since Class A provides lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is device that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channel at the same time. Gateways are typically connected to power grid so power consumption of a gateway is insignificant most of the deployments.

\subsection{Network server}
\par LoRaWAN network server (NS) is server that provides MAC layer processing. Network server routes messages from application to end nodes, and vice versa. Network server can be used for tweaking end node parameters like channel, transmit power, SF to increase network performance.

\section{LoRa}
\par LoRa is the name of the proprietary physical layer radio and chipset technology that provides wireless link solution for low power wide area networks. LoRa is designed and patented by Semtech Corporation. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). Chirp is a sinusoidal signal which its  frequency increases over time. The chirp frequency increases linearly and sweeps entire bandwidth. We invite readers to refer to \cite{AN1200.22} for further details.

\subsection{Spreading Factor}
\par Spreading factor (SF) is the ratio between symbol rate and chip rate. The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}. SF can take values between 7 to 12. SF also determines data rate of a LoRa transmission \cite{AN1200.22}.

\begin{equation} \label{eq:bit_rate_sf}
R_{b}|_{bps} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW|_{Hz}} \right]}
\end{equation}

Where, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

\par While BW and CR are constant, as the SF increases the data rate decreases. Increasing SF makes the signal more resilient to noise thus increases the transmission range. Increasing SF also increases transmission duration which increases power consumption. Therefore it is possible to trade between data rate, range and power consumption by  changing SF.

@@

\par Initially, a node is not aware of how far it is from a GW. However, a node can guess how far away from the gateway by observing received signal power of a downlink transmission. If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption. Also GW can request from a node to decrease node's SF or transmit power. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.

\par In Figure \ref{fig:collision}, a LoRaWAN network distributed around a single GW is illustrated. Different color rings represent achievable range of different SF from GW and different color circles represent nodes with selected SF. Most of the end devices which close to the GW will fall into lowest SF (SF7) area section. End devices which close to the GW will probably select lowest SF all the time. Which results a lot of collisions between same SF transmissions. Collision number increases while number of end devices close to the GW increases.

@@

\section{Other Related Works} \label{Other Related Works}
The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately. We listed some of these works which studying LoRaWAN and LoRa SF.

\par In \cite{7996384}, the authors evaluated the performance of LoRa networks in a smart city scenario. The
authors proposed a link measurement and a link performance model for LoRa. The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions. They implement a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks. Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases. They also show that SF assignment has great effect on network performance.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements. Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.

\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal. Their experimental results show that this power difference is around 16 dB and this power difference can be seen when  interferer is close to receiver or multiple interferer can create this power difference cumulatively.

\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions using the same SF as well as different SFs. They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage performance with respect to the distance from the gateway for modeling interference in multiple gateway scenarios. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even if they can able to communicate with lower SF. This may result lower collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, how GW assign SF to which node becomes an another issue. Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions. In multiple GW scenarios, this approach may increase the collision number with nodes in other GW's range. Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules. For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collision number.  This technique learns nodes behavior of a network. A NS can keep track of successful uplink transmission from nodes and their SFs. NS can also keep track of some of the collided transmission if header part of the packet does not interfered at GW. However, NS cannot keep track of transmissions with receive power lower than GW sensitivity. Using these obtained information NS can train a classifier for predicting transmission result for specific node and specific SF input. Using this predictor GWs can assign the lowest possible SF to nodes considering the collision probability.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality. The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorafs. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed random circle shape network topology with radius, number of nodes and number of gateways inputs. Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method. With these input, the simulator provide outputs of total number of generated packets, number of successfully received packet, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).

@@

\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}. In our simulator we assume that nodes are always uses maximum allowed transmit power which is 14 dBm. Different channel transmissions are independent from each other. But, in this work we focuse on SF orthogonality. Thus, we only utilized single channel transmissions in our simulator.

@@

\begin{equation} \label{eq:transmission_duration}
t_{x}|_{s} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}
\end{equation}

\subsection{Machine Learning Model}
\par We have been able to generate mass amount of LoRaWAN transmission data for different topologies using our  tool. To come up with a solution for SF assignment issue, we have integrated a machine learning scheme to our simulator. This will allow us train a model from generated transmission data and we use this trained model to select best possible SF for nodes using prediction of the model. Tool first run a simulation with assigning random SF to nodes. Simulation results are combined into four columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data is fed to Python scikit-learn decision tree classifier. After the model is created, a second simulation is run with a predictor. Model tries to find the lowest possible SF with prediction of successful transmission.

@@

\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning predictor SF scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Our predictor model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation. Predictor scheme gives better PDR when number of nodes increases since number of interference increases when number of nodes increases. Predictor scheme can improve performance when LoRa interference is present. Also predictor scheme gives better results when nodes are deployed close to the gateway, since nodes have a margin to increase their SFs. If a node is far away from the GW, then predictor scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of different prediction SF assignment scheme.}
\label{fig:predict_pdr}
\end{figure}

>>>>

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas. Focus of this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa modulation trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance. In this work, a simulation environment to study different SF assignment schemes is implemented. Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization. Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on growth of number of IoT applications. LPWAN technologies addresses some of the well-known wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication technologies (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications do not require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox are known for ultra low power consumption and resilience to interference.  While NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. LoRaWAN networks can be deployed as a private network like WiFi. However, Sigfox and NB-IoT are only available with operator contract \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size. Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

@@

\section{LoRaWAN} \label{LoRaWAN}
\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently yet incorrectly used for each other.

\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and it is maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN. LoRa can be used as a wireless link technology without complying LoRaWAN, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides inter-operability between different LoRa networks. LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.  LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is a low power embedded device that only communicates between gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C. Different classes provides LPWAN solutions to different deployments and topologies. Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateways. Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission. This provides Class C end nodes to low latency donwlink communication but it requires more power consumption. In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.

\subsection{Gateway}
\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes. A typical gateway can receive from multiple channels at the same time. Gateways are typically connected to power grid so power consumption of a gateway is insignificant in most of the deployments.

\subsection{Network server}
\par LoRaWAN network server (NS) is a server that provides MAC layer processing. Network server routes messages from application to end nodes and vice versa. Network server can be used for tweaking end node parameters like channel, transmit power and SF to increase network performance.

\section{LoRa}
\par LoRa is the name of the proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks. LoRa is designed and patented by Semtech Corporation. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). A chirp is a sinusoidal signal which its frequency increases over time. Chirp frequency increases linearly and sweeps the entire bandwidth. We invite readers to refer to \cite{AN1200.22} for further details.

\subsection{Spreading Factor}
\par Spreading factor (SF) is the ratio between symbol rate and chirp rate. The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}. SF can take values between 7 to 12. SF also determines data rate of a LoRa transmission \cite{AN1200.22}.

\begin{equation} \label{eq:bit_rate_sf}
R_{b} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW|_{Hz}} \right]} \ bps
\end{equation}

Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.

\par When BW and CR are constant, as the SF increases, the data rate decreases. Increasing the SF makes the signal more resilient to noise thus increases the transmission range. Increasing the SF also increases the transmission duration which increases the power consumption. Therefore it is possible to trade between data rate, range and power consumption by changing SF.

@@

\par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW. However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission. If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption. This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper. Also a GW can request from a node to decrease node's SF or transmit power.

\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated. Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes. Most of the end devices which close to the GW will fall into lowest SF (SF7) area section. End devices which close to the GW will probably select lowest SF all the time. Which results a lot of collisions between same SF transmissions. Number of collisions increases while number of end devices close to the GW increases.

@@

\section{Other Related Works} \label{Other Related Works}
The literature related to the work presented in this paper has started growing recently. LPWAN technologies and especially LoRa  attracted researchers attentions lately. We listed some of these works which studies LoRaWAN and LoRa SF.

\par In \cite{7996384}, the authors evaluated the performance of LoRaWAN networks in a smart city scenario. The authors proposed a link measurement and a link performance model for LoRa. The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions. They implemented a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks. Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases. They also show that SF assignment has great effect on LoRaWAN network performance.

@@

\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements. They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements. Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.

\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal. Their experimental results show that this power difference is around 16 dB and this power difference can be seen when a interferer is close to receiver or multiple interferer can create this power difference cumulatively.

\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with same and different SFs. They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage  distance from the gateway for modeling interference in multiple gateway scenarios. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

\section{Proposed Technique} \label{Proposed Technique}
\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even they can able to communicate with lower SF. This may result lower collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}. Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}. However, how GW should assign SF to which node becomes an another issue. Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions. In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range. Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.

\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules. For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions.  This technique learns nodes behavior of a network. A NS can keep track of successful uplink transmission from nodes and their SFs. NS can also keep track of some of the collided transmission if header part of the packet does not interfered at GW. However, NS cannot keep track of transmissions with receive power lower than GW sensitivity. Using these obtained information NS can train a classifier for predicting transmission result for specific node and specific SF input. Using this prediction model GWs can assign the lowest possible SF to nodes considering the collision probability.

@@

\section{Simulation Environment} \label{Simulation Environment}
\par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality. The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorasf. Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies. Simulator can generate uniformly distributed random circle shape network topology with radius, number of nodes and number of gateways inputs. Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method. With these input, the simulator provide outputs of total number of generated packets, number of successfully received packet, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).

@@

\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}. In our simulator we assume that nodes are always uses maximum allowed transmit power which is 14 dBm. Different channel transmissions are independent from each other. But, in this work we focus on SF orthogonality. Thus, we only utilized single channel transmissions in our simulator.

@@

\begin{equation} \label{eq:transmission_duration}
t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
\end{equation}

\subsection{Machine Learning Model}
\par We have been able to generate mass amount of LoRaWAN transmission data for different topologies using our  tool. To come up with a solution for SF assignment issue, we have integrated a machine learning scheme to our simulator. This will allow us train a model from generated transmission data and we use this trained model to select best possible SF for nodes using prediction of the model. Tool first run a simulation with assigning random SF to nodes. Simulation results are combined into four columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission. Transmission result can be successful, interfered or under sensitivity. These data is fed to Python scikit-learn decision tree classifier. After the model is created, a second simulation is run with a prediction scheme. Model tries to find the lowest possible SF with prediction of successful transmission.

@@

\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown. Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second. Our prediction model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation. Predictor scheme gives better PDR when number of nodes increases since number of interference increases when number of nodes increases. Predictor scheme can improve performance when LoRa interference is present. Also predictor scheme gives better results when nodes are deployed close to the gateway, since nodes have a margin to increase their SFs. If a node is far away from the GW, then predictor scheme cannot increase the SF to avoid interference since the SF is already high.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
\caption{PDR of prediction SF assignment scheme.}
\label{fig:predict_pdr}
\end{figure}
